---
title: "NYC_Kaggle"
author: "Chris"
date: "June 28, 2018"
output: html_document
---
Start deriving insights
Instead of more coding. Let's first do more research on these schools
The eastern schools have a high black/hispanic however they seem to have very good income
and economic need ratio. whats the difference?
The southern and middle have the high Scores on the exams
Community schools are way more prevalent in our black/hispanic areas
Not suprisingly our areas also some the majority of the low # of SHSAT takers
Why are Glendale and Elmhurst so high in test takers?
Why is Broad Channel so low?
Make sure to ignore school for the performing arts i think?
Where are these higher acheiving schools located?

First lets read in the data
```{r}

library(readr)
exploration_data2016 <- read_csv("~/Kaggle Data Sets/Data Science for Good/2016 School Explorer.csv", 
    na = "N/A")
View(exploration_data2016)
SHSAT_Stats <- read_csv("~/Kaggle Data Sets/Data Science for Good/D5 SHSAT Registrations and Testers.csv")
View(SHSAT_Stats)
```

There are too many variables to use head or str on. Lets start with some mappings.
First up is Economic Need Index
```{r}
library(leaflet)
library(dplyr)
exploration_data2016[["School_Economic_Bin"]] <- cut(exploration_data2016[["Economic Need Index"]], 
                                                   breaks=c(quantile(exploration_data2016[["Economic Need Index"]], probs = seq(0, 1, by = 0.25), na.rm=T)))

factpal <- colorFactor(c("green","blue","yellow","red"), 
                       exploration_data2016[["School_Economic_Bin"]])

exploration_data2016 %>% filter(!is.na(School_Economic_Bin))  %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(School_Economic_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~School_Economic_Bin,
    title = "Economic Need Index") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```


```{r}
exploration_data2016[["School Income Estimate"]] <- parse_number(exploration_data2016[["School Income Estimate"]])
exploration_data2016[["School_Income_Bin"]] <- cut(exploration_data2016[["School Income Estimate"]], 
                                                   breaks=c(quantile(exploration_data2016[["School Income Estimate"]], probs = seq(0, 1, by = 0.25), na.rm=T)))

factpal <- colorFactor(c("red","yellow","blue","green"), 
                       exploration_data2016[["School_Income_Bin"]])


exploration_data2016 %>% filter(!is.na(School_Income_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(`School_Income_Bin`)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~`School_Income_Bin`,
    title = "School Income Estimate") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```
Seems the bronx are in need the most which is expected

Lets look at black and hispanic populations
```{r}
exploration_data2016[["Percent Black"]] <- parse_number(exploration_data2016[["Percent Black"]])
exploration_data2016[["Black_Bin"]] <- cut(exploration_data2016[["Percent Black"]], 
                                                   breaks=c(0,25,50,75,100))

factpal <- colorFactor(c("green","blue","yellow","red"), 
                       exploration_data2016[["Black_Bin"]])


exploration_data2016 %>% filter(!is.na(Black_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(`Black_Bin`)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~`Black_Bin`,
    title = "Black Percent") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

```{r}
exploration_data2016[["Percent Hispanic"]] <- parse_number(exploration_data2016[["Percent Hispanic"]])
exploration_data2016[["Hispanic_Bin"]] <- cut(exploration_data2016[["Percent Hispanic"]], 
                                                   breaks=c(0,25,50,75,100))

factpal <- colorFactor(c("green","blue","yellow","red"), 
                       exploration_data2016[["Hispanic_Bin"]])


exploration_data2016 %>% filter(!is.na(Hispanic_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(Hispanic_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~Hispanic_Bin,
    title = "Hispanic Percent") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

```{r}
exploration_data2016[["Percent Black / Hispanic"]] <- parse_number(exploration_data2016[["Percent Black / Hispanic"]])
exploration_data2016[["Both_Bin"]] <- cut(exploration_data2016[["Percent Black / Hispanic"]], 
                                                   breaks=c(0,25,50,75,100))

factpal <- colorFactor(c("green","blue","yellow","red"), 
                       exploration_data2016[["Both_Bin"]])


exploration_data2016 %>% filter(!is.na(Both_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(Both_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~Both_Bin,
    title = "Hispanic Percent") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

```{r}
exploration_data2016[["Percent ELL"]] <- parse_number(exploration_data2016[["Percent ELL"]])
exploration_data2016[["ELL_Bin"]] <- cut(exploration_data2016[["Percent ELL"]], 
                                                   breaks=c(0,25,50,75,100))

factpal <- colorFactor(c("green","blue","yellow","red"), 
                       exploration_data2016[["ELL_Bin"]])


exploration_data2016 %>% filter(!is.na(ELL_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(ELL_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~ELL_Bin,
    title = "ELL Percent") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```


```{r}
exploration_data2016[["Student Attendance Rate"]] <- parse_number(exploration_data2016[["Student Attendance Rate"]])
exploration_data2016[["Attendance_Bin"]] <- cut(exploration_data2016[["Student Attendance Rate"]], 
                                                   breaks=c(0,75,85,95,100))

factpal <- colorFactor(c("red","yellow","blue","green"), 
                       exploration_data2016[["Attendance_Bin"]])


exploration_data2016 %>% filter(!is.na(`Student Attendance Rate`)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(Attendance_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~Attendance_Bin,
    title = "ELL Percent") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

```{r}
exploration_data2016[["Effective School Leadership %"]] <- parse_number(exploration_data2016[["Effective School Leadership %"]])
exploration_data2016[["Leadership_Bin"]] <- cut(exploration_data2016[["Effective School Leadership %"]], 
                                                   breaks=c(0,75,85,95,100))

factpal <- colorFactor(c("green","blue","yellow","red"), 
                       exploration_data2016[["Leadership_Bin"]])


exploration_data2016 %>% filter(!is.na(Leadership_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(Leadership_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~Leadership_Bin,
    title = "Leadership Percent") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

```{r}
exploration_data2016[["Average ELA Proficiency"]] <- parse_number(exploration_data2016[["Average ELA Proficiency"]])
exploration_data2016[["ELA_Bin"]] <- cut(exploration_data2016[["Average ELA Proficiency"]], 
                                                   breaks=c(0,2,2.67,3.33,4))

factpal <- colorFactor(c("red","yellow","blue","green"), 
                       exploration_data2016[["ELA_Bin"]])


exploration_data2016 %>% filter(!is.na(ELA_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(ELA_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~ELA_Bin,
    title = "ELA Average") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```


```{r}
exploration_data2016[["Average Math Proficiency"]] <- parse_number(exploration_data2016[["Average Math Proficiency"]])
exploration_data2016[["Math_Bin"]] <- cut(exploration_data2016[["Average Math Proficiency"]], 
                                                   breaks=c(0,2,2.67,3.33,4))

factpal <- colorFactor(c("red","yellow","blue","green"), 
                       exploration_data2016[["Math_Bin"]])


exploration_data2016 %>% filter(!is.na(Math_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(Math_Bin)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~Math_Bin,
    title = "Math Average") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

```{r}

factpal <- colorFactor(c("green", "red"), 
                       exploration_data2016[["Community School?"]])


exploration_data2016  %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~factpal(`Community School?`)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~`Community School?`,
    title = "Community School") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

```{r}
factpal <- colorFactor(topo.colors(11), 
                      exploration_data2016[["Grade High"]])


exploration_data2016  %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, color = ~factpal(`Grade High`)) %>% 
  addTiles() %>%
  addLegend("bottomright", pal = factpal, values = ~`Grade High`,
    title = "Highest Grade") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```


Get data around the all student columns
```{r}
#sum across multiple
total_number <- exploration_data2016 %>% select(Latitude, Longitude, District, contains("All Students Tested")) %>%
  mutate(row_sums = rowSums(.[4:15])) 

total_number[["Total_Takers_Bin"]] <- cut(total_number[["row_sums"]], 
                                                   breaks=c(0,500,1000,2000,5000))

qpal <- colorFactor(c("red","yellow","blue","green"), 
                       total_number[["Total_Takers_Bin"]])

total_number %>% filter(!is.na(Total_Takers_Bin)) %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = ~qpal(Total_Takers_Bin)) %>% 
  addTiles() %>%
    addLegend("bottomright", pal = qpal, values = ~Total_Takers_Bin,
    title = "# of Exam Takers") %>%
  fitBounds(min(exploration_data2016$Longitude), min(exploration_data2016$Latitude),
            max(exploration_data2016$Longitude), max(exploration_data2016$Latitude))
```

get a subset of columns for per city of only rows with min at least grade 3 and max grade 8
```{r}
library(ggplot2)
library(tidyr)
library(stringr)
library(forcats)
time_series <- exploration_data2016 %>% filter(`Grade Low` %in% c("03", '02', '01', 'PK', '0K') & 
                                  `Grade High` %in% c("08")) %>% select(City, contains("ELA - All Students Tested")) %>% group_by(City) %>%
                      summarise_all(mean) %>% gather(Test_Year, Average_Taken, `Grade 3 ELA - All Students Tested`:`Grade 8 ELA - All Students Tested`) 

time_series$Test_Year <- parse_number(time_series[["Test_Year"]])



g <- ggplot(time_series, aes(x=Test_Year, y = Average_Taken)) + geom_line(aes(color=City))
library(plotly)
ggplotly(g)
```

Make a heatmap next of the SHSHAT

```{r}
SHSAT_Stats["Percent_Registration"] <- SHSAT_Stats["Number of students who registered for the SHSAT"]/SHSAT_Stats["Enrollment on 10/31"]
SHSAT_Stats["Percent_Taken_OfReg"] <- SHSAT_Stats["Number of students who took the SHSAT"]/SHSAT_Stats["Enrollment on 10/31"]
spread_data <- SHSAT_Stats %>% spread(`Year of SHST`, value = `Percent_Registration`, fill = 0) %>% select(`School name`, `2013`, `2014`, `2015`, `2016`)

spread_agg <- spread_data %>% group_by(`School name`) %>% summarise(sum_2013 = sum(`2013`),
                                                                    sum_2014 = sum(`2014`),
                                                                    sum_2015 = sum(`2015`),
                                                                    sum_2016 = sum(`2016`))
library(RColorBrewer)
plot_ly(
  x = c("2013", "2014", "2015", "2016"), y = spread_agg[["School name"]],
        z = as.matrix(spread_agg[2:5]), type = "heatmap"
)

```

```{r}
spread_data <- SHSAT_Stats %>% spread(`Year of SHST`, value = `Percent_Taken_OfReg`, fill = 0) %>% select(`School name`, `2013`, `2014`, `2015`, `2016`)

spread_agg <- spread_data %>% group_by(`School name`) %>% summarise(sum_2013 = sum(`2013`),
                                                                    sum_2014 = sum(`2014`),
                                                                    sum_2015 = sum(`2015`),
                                                                    sum_2016 = sum(`2016`))
plot_ly(
  x = c("2013", "2014", "2015", "2016"), y = spread_agg[["School name"]],
        z = as.matrix(spread_agg[2:5]), type = "heatmap"
)
```








After analysis. Lets first isolate the data with just the rows with > 75% black/hispanic
```{r}
heavy_ethnic <-  exploration_data2016 %>% filter(`Percent Black / Hispanic` >= 75)
heavy_ethnic  %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = "Red") %>% 
  addTiles() %>%
  fitBounds(min(heavy_ethnic$Longitude), min(heavy_ethnic$Latitude),
            max(heavy_ethnic$Longitude), max(heavy_ethnic$Latitude))
```
From this, we see there are more or less 4 main parts to initially focus our attention on (low hanging fruit first). The bronx area, brooklyn, east queens and lower manhattan.


Lets see how many of these areas have good math and ELA scores
```{r}
high_learners <-  heavy_ethnic %>% filter(`Average ELA Proficiency` >= 3 & `Average Math Proficiency` >= 3)
high_learners  %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = "Red") %>% 
  addTiles() %>%
  fitBounds(min(heavy_ethnic$Longitude), min(heavy_ethnic$Latitude),
            max(heavy_ethnic$Longitude), max(heavy_ethnic$Latitude))
```
How does this compare to non ethnic places. Manhattan does really well. that might be a good area to look at.
```{r}
high_learn <-  exploration_data2016 %>% filter(`Percent Black / Hispanic` < 75 & `Average ELA Proficiency` >= 3 & `Average Math Proficiency` >= 3)
high_learn  %>% leaflet() %>% addCircles(lng = ~Longitude, lat = ~Latitude, 
                                             color = "Red") %>% 
  addTiles() %>%
  fitBounds(min(heavy_ethnic$Longitude), min(heavy_ethnic$Latitude),
            max(heavy_ethnic$Longitude), max(heavy_ethnic$Latitude))
```

